Untitled

### 2020-12-18 ms/manuscript.Rmd
Update MS

```{r load-data}
loadd(
  list = c(
    "climate",
    "filmy_species",
    "moorea_sites",
    "phylosig",
    "range_types",
    "combined_species_means",
    "t_test_results"
  )
)
```

@@

Filmy ferns (Hymenophyllaceae) represent a particularly interesting case for studying the ecological and evolutionary significance of vegetative DT across life stages in ferns.
Filmy ferns are a large family (ca. 430 spp.; PPG I, 2016) of primarily tropical, leptosporangiate ferns named for their remarkably thin leaf laminae, which are usually only a single cell layer in thickness and lack cuticle or stomata (Fig. 1A--D).
Filmy fern sporophytes are an exception to the 'typical' fern sporophyte described above because they have reverted to poikilohydry: their thin laminae allow for passive, rapid uptake and loss of moisture, thereby maintaining nearly constant equilibrium with atmospheric humidity [@Raven1999].
Many species of epiphytic filmy ferns have extremely reduced root systems [@Schneider2000], relying completely upon absorption through leaf laminae for water [@Shreve1911], although terrestrial species possess true roots and use ground water [@Dubuisson2011].
As would be expected from their poikilohydric nature, DT of varying degrees has been reported in sporophytes of many filmy fern species (see references below).

@@

We used microclimate data collected by @Nitta2020.
This dataset includes relative humidity (RH) and temperature measured every 15 min from `r climate_start` to `r climate_end` at the sampling sites with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA).
The dataloggers were outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation, and placed in pairs at each site (one 'epiphytic' datalogger at ca. 1.5 m on a tree, one 'terrestrial' datalogger at \circa{ }10 cm next to the tree).
During the study period, some dataloggers malfunctioned, presumably due to prolonged exposure to high humidity.
Any days during which one or more dataloggers failed are excluded from the dataset.
One terrestrial datalogger (Mt. Tohiea 393 m site) was lacking data for more than half of the survey period, but had data available for the same period from a preliminary survey conducted during the previous year (2012--2013), so the previous year's data for this site was used instead.
The dataset includes `r climate_n_days` days of data from `r climate_n_sites` sites (`r climate_n_loggers` dataloggers).

@@

Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8--16 individuals (one individual = single whole frond including a ca. 2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment chlorophyll fluorescence (F~v~/F~m~) was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100% RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22°C to 25°C.
Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (-282 MPa), Mg(NO~3~)~2~ (-86 MPa), and NaCl (-38 MPa).
Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible; we therefore used the same treatment for all gametophytes (2 d at -86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described below.
Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
Following the desiccation treatment, plants were rewetted using water from a local stream.
Plants were allowed to rehydrate, and chlorophyll fluorescence was again measured at 0.5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment chlorophyll fluorescence yield [@WatkinsJr.2007c].

@@

In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes (Table 1) including *Callistopteris apiifolia* (C. Presl) Copel., *Crepidomanes minutum* var. 2 and 3, *Hymenophyllum digitatum* (Sw.) Fosberg, *Hymenophyllum javanicum* Spreng., *Hymenophyllum pallidum* (Blume) Ebihara & K.
Iwats., and *Polyphlebium borbonicum* (Bosch) Ebihara & Dubuisson.
All of these have sporophytes that are normally only found in cloud forest (ca. 500--1200 m), but gametophytes that are distributed to lower elevations.

Sporophytes of two taxa, *C. minutum* var. 3 and *Hymenophyllum braithwaitei* (Ebihara & K. Iwats.), did not occur in the study plots, but each was observed from a single population on Mt. Tohiea, at ca. 1000 m and near the summit (1206 m), respectively.
We include these observations for comparison of ranges between sporophytes and gametophytes.

@@

Gametophytes of low elevation epiphytes had the highest DT (ca. 80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from ca. 25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; Fig. 5).
DT of gametophytes did not significantly exceed that of sporophytes for any species except for saxicolous *P. endlicherianum* (Fig. 5; *t*-test, *P* = `r ttest_pval("Polyphlebium_endlicherianum", "recover")`), and was significantly lower in gametophytes than sporophytes of some high elevation epiphytic species including *C. minutum* var. 2, *H.* *digitatum*, *H. polyanthos*, and *P. borbonicum* (Fig. 5; *t*-test, *P* = `r ttest_pval("Crepidomanes_minutum2", "recover")`, `r ttest_pval("Hymenophyllum_digitatum", "recover")`, `r ttest_pval("Hymenophyllum_polyanthos", "recover")`, and `r ttest_pval("Polyphlebium_borbonicum", "recover")`, respectively).

@@

Desiccation tolerance tended to increase with VPD for both gametophytes and sporophytes (Fig. 6).
These trends were significant when analyzed using PGLS for sporophytes (*P* = 0.036) but not gametophytes (*P* = 0.123; Appendix S5).

@@

We found a wide variation of DT in filmy fern sporophytes on Moorea, from no tolerance in terrestrial species to tolerance of water potentials well below --200 MPa in some epiphytes.
Levels of similarly extreme DT have also been reported in other species of *Hymenophyllum* (Proctor, 2003, 2012) and a general correlation of habitat with DT in filmy ferns has been previously established (Shreve, 1911; Proctor, 2003, 2012; Parra et al., 2009; Saldan˜a et al., 2014; Testo et al., unpublished), but no other studies to our knowledge have documented a similar range of DT values in filmy ferns from a single site.
One may wonder then, what is the adaptive significance of being able to withstand drying at --200 MPa vs. --100 MPa when the ambient water potential is typically *\>*--5 MPa and only rarely, if ever, drops below −50 MPa.
The recent finding of a constitutive mechanism for DT in filmy ferns (Cea et al., 2014) supports the need for a relatively wide 'safety margin,' and similar (or even more extreme) levels of constitutive DT are well known in bryophytes (Oliver et al., 1993).
The conservative behavior of fern stomata, which lack a response to the plant stress hormone ABA, is well documented (Brodribb and McAdam, 2011; McAdam and Brodribb, 2012, 2013).
It is possible that filmy ferns also behave conservatively with respect to water-stress.
We suggest that the variation in DT observed here is relevant because it is correlated with habitat, and likely reflects overall physiological tolerance, not just artificial extremes.
For example, sporophytes of some high elevation epiphytes that we surveyed (*C. minutum* var. 2, *H. digitatum*, *H. multifidum*, *H. polyanthos*) seem to be able to better withstand intense vs. moderate desiccation, with higher recovery values at --282 and --86 MPa vs. --38 MPa when dry for 15 d.
This may be an adaptation to cloud forest conditions; although cloud forests are frequently moist due to constant fog, dry periods are intense during the times when there is no cloud cover.
The complete lack of DT in terrestrial species is not surprising, considering they have much more direct access to water than epiphytes and therefore substantially less selective pressure to maintain DT.
Since controls kept at 100% RH maintained F~v~/F~m~ values near pre-treatment levels in all sporophytes (Appendix S3), it is unlikely that the lack of tolerance seen in terrestrial species is due solely to removal from the soil.
Comparison of DT with the minimum water potential envelope of each species (Fig. 6) indicates that sporophytes of non-tolerant terrestrial and saxicolous species may not be able to occur in habitats with minimum water potential exceeding ca.
--7 MPa.
In contrast, habitats with minimum water potential below --7 MPa are nearly all epiphytic, and the filmy fern sporophytes occurring there are all capable of DT.

<!--

@@

Contrary to our hypotheses, gametophytes of filmy ferns are in general not more tolerant of the abiotic environment (light levels and water potential) than sporophytes.
This differs from other fern groups with desiccation tolerant gametophytes but desiccation sensitive sporophytes (Watkins et al., 2007).
In the filmy ferns we studied, DT levels of both sporophytes and gametophytes largely reflected habitat; in general, epiphytic species tended to have higher DT than terrestrial or saxicolous species, regardless of life stage.
Interestingly, this echoes much earlier field observations (not experimental data) by Holloway (1930) on filmy fern gametophytes in New Zealand: he reported apparent DT in gametophytes of *Hymenophyllum rarum* R.
Br.
and *H. villosum* Colenso, both high elevation epiphytes with desiccation tolerant sporophytes, and lack of DT in terrestrial filamentous gametophytes of *Polyphlebium colensoi* (Hook.) Ebihara & K.
Iwats., *Abrodictyum strictum* (Menzies ex Hook. & Grev.) Ebihara & K.
Iwats., and *A. elongatum* (A. Cunn.) Ebihara & K.
Iwats.
occupying moist, protected sites on the forest floor.
Watkins et al. (2007) also found that DT of fern gametophytes tended to match habitat, with terrestrial species having lower DT than epiphytes.
However, we found that within high elevation epiphytes, gametophytes had lower DT than sporophytes, and only one species (saxicolous *P. endlicherianum*) with gametophytes that had significantly greater DT than sporophytes (Fig. 5).
At high elevation sites (i.e., cloud forest), epiphytic ferns often occur in a dense matrix of bryophytes (Fig. 1B).
It is possible that this bryophyte cover helps to retain moisture during brief periods of drought and thus protect the epiphytic fern gametophytes that grow amongst them.
More detailed studies of the high elevation epiphytic environment and interactions between co-occurring gametophytes of ferns and bryophytes are needed to test this hypothesis.

Although we predicted that species with gametophytes that occur beyond the range of their sporophytes would have higher DT and be adapted for higher light levels, this was not the case.
This is perhaps best illustrated by *C. apiifolia* (Fig. 1D, E, H).
Sporophytes of this species are usually restricted to moist cloud forest sites from ca.
600--1200 m, but we observed gametophytes over a much wider range, down to 200 m.
A similar distribution pattern has been reported in congeneric *Callistopteris baldwinii* (D. C. Eaton) Copel.
from Hawaii (Dassler and Farrar, 1997).
Since gametophytes of *C. apiifolia* are common on Moorea and have a unique morphology (Nitta et al., 2016), we were able to collect enough material of this species for additional tests of DT at higher humidity levels (48 h at 75% and 100% RH; results not shown).
However, it still failed to recover from even these gentle treatments.
This suggests that the widespread gametophytes of *C. apiifolia* do not rely on DT to attain their distribution, but instead must do so via some other mechanism.
Although we measured terrestrial microhabitats by positioning sensors a few cm above the ground, these sensors are still outside of the boundary layer.
It is likely that the gametophytes of *C. apiifolia* do not actually experience the conditions measured with our 'terrestrial' dataloggers, but instead grow within a much more consistently moist boundary layer.
Gametophytes of *C. apiifolia* produce gemmae (Fig. 1E, H), and often form dense clonal mats.
In the case that their microhabitats do become too dry, they may suffer loss at the edge of the population, then recover by clonal growth (Farrar et al., 2008).
Furthermore, the complex surface of clonal mats has been shown to increase the boundary layer and decrease evaporation rates in mosses (Rice et al., 2001), and may function similarly in *C. apiifolia*.
Other filmy fern species also produce gemmae, but they are much more frequent in *C. apiifolia*, and the large size of the clonal mats formed by *C. apiifolia* indicate that it may have a high growth rate (J. Nitta, pers. obs.).
It is possible that there is a tradeoff between growth rate and DT, such that species capable of rapid asexual growth like *C. apiifolia* are unable to tolerate desiccation, but slower growing epiphytes such as *Hymenophyllum* spp.
can better tolerate it (Oliver et al., 2000).
Growth rates of tolerant and non-tolerant gametophytes including quantification of gemmae production should be investigated in these species to test this hypothesis.

@@

Unlike DT, photosynthetic parameters did not vary significantly between species from different habitats.
Rather, we observed a clear intergenerational difference, with gametophytes consistently adapted for lower light levels than sporophytes (Fig. 5, Table 3).
In the only other study to our knowledge that compared photosynthetic rates between gametophytes and sporophytes of a filmy fern, Johnson et al. (2000) also found that gametophytes of *V. speciosa* reached maximum ETR at lower levels than sporophytes: in gametophytes of this species, photosynthesis became saturated at a PPFD of 4--5 µmol·m^--2^ ·s^--1^ in the field or 30--50 µmol·m^--2^·s^--1^ in the lab vs. 30--50 µmol·m^--2^ ·s^--1^ for the sporophyte (lab measurement only).
Although we found similar differences in direction between gametophytes and sporophytes, the measurements of Johnson et al. (2000) are considerably lower than what we observed in the field for both Moorean filmy fern gametophytes (PPFD~95%~ ca. 40--60 µmol·m^--2^ ·s^--1^) and sporophytes (PPFD~95%~ ca. 130--160 µmol·m^--2^ ·s^--1^).
This makes sense in light of the fact that *V. speciosa* is an extremely deep-shade plant, with gametophytes occupying protected rock crevices that receive less than 1 µmol·m^--2^·s^--1^ PPFD for most of the day (Johnson et al., 2000); many of the gametophytes sampled in our study occur in more exposed conditions.
It is surprising, however, that we did not observe any differences between species in different habitats, given that light levels should vary between tree trunks and the forest floor, at least on a ca.
1 m scale.
It is possible that the actual light levels experienced by gametophytes growing in these habitats may not vary significantly if they occupy protected microsites such as bryophyte mats or crevices in rock or bark.

@@

Previous studies of filmy fern ecophysiology also found a link between habitat and DT and/or photosynthetic rates (Proctor, 2003, 2012; Parra et al., 2009; Saldan˜a et al., 2014); however, ours is the first to our knowledge to investigate this pattern while taking phylogeny into account.
We detected significant phylogenetic signal in DT, suggesting that phylogeny should be considered when interpreting differences in DT between species (Table 2).
In contrast, the lack of phylogenetic signal in photosynthetic traits, coupled with the clear intergenerational differences we observed, suggest that sporophytes and gametophytes are adapted for different light levels regardless of clade or ecological niche.
In a more densely sampled study of morphological evolution in the trichomanoid clade, Dubuisson et al. (2013) identified several morphological changes including reduced stele and root systems associated with the evolution of epiphytic growth within the HE subclade (comprising the genera *Polyphlebium*, *Didymoglossum*, *Crepidomanes*, and *Vandenboschia*).
It is possible that both reduced morphology and increased DT are parts of an integrated adaptive strategy for epiphytic growth within this subclade of filmy ferns, and have evolved together repeatedly.
However, Dubuisson et al. (2013) were unable to determine if these morphological changes had occurred independently, or if there was a single origin at the base of the HE clade followed by multiple losses.
Increased sampling of both morphological and physiological traits of additional trichomanoid taxa is needed to distinguish between these scenarios.

@@

In filmy ferns, both sporophyte and gametophyte generations are poikilohydric; however, we have found here that this does not mean they necessarily share similar physiological optima.
Although filmy fern sporophytes are often of small stature, elevating a leaf even a few cm above the substrate surface means escape from the boundary layer and exposure to substantially greater evapotranspiration.
It is likely that even in this clade, wherein the physiological divide between sporophyte and gametophyte is probably the smallest amongst ferns, the two generations experience substantially different selective pressures.
Fossils indicate that some of the earliest vascular plants may have had sporophytes and gametophytes that were both branched and small (Kenrick and Crane, 1997), but it is unclear at what point selective pressures began to diverge to favor large, complex sporophytes and small, simple gametophytes.
The relevance of plant size to water relations is supported by a recent study showing increased DT in smaller size classes of the epiphytic fern *Asplenium auritum* Sw.
(Testo and Watkins, 2012).
Future studies of DT in filmy ferns should include juvenile stages of the sporophyte, which are expected to experience similar microhabitats as gametophytes, to better determine the role of microhabitat vs. life stage (and plant size) on DT.
Finally, more detailed analyses of the strategies used by each generation to cope with desiccation stress, such as transcriptomics, are needed to better understand the ecology and evolution of filmy ferns as well as the transition from desiccation tolerance to avoidance in vascular plants.

@@

--\>

@@

<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
<div id="refs"></div>

>>>>

```{r load-data}
loadd(
  list = c(
    "climate",
    "filmy_species",
    "moorea_sites",
    "phylosig",
    "range_types",
    "combined_species_means",
    "t_test_results",
    "env_range_dt_model_summary"
  )
)
```

@@

Filmy ferns (Hymenophyllaceae) represent a particularly interesting case for studying the ecological and evolutionary significance of vegetative DT across life stages in ferns.
Filmy ferns are a large family (\circa{ }430 spp.; PPG I, 2016) of primarily tropical, leptosporangiate ferns named for their remarkably thin leaf laminae, which are usually only a single cell layer in thickness and lack cuticle or stomata (Fig. 1A--D).
Filmy fern sporophytes are an exception to the 'typical' fern sporophyte described above because they have reverted to poikilohydry: their thin laminae allow for passive, rapid uptake and loss of moisture, thereby maintaining nearly constant equilibrium with atmospheric humidity [@Raven1999].
Many species of epiphytic filmy ferns have extremely reduced root systems [@Schneider2000], relying completely upon absorption through leaf laminae for water [@Shreve1911], although terrestrial species possess true roots and use ground water [@Dubuisson2011].
As would be expected from their poikilohydric nature, DT of varying degrees has been reported in sporophytes of many filmy fern species (see references below).

@@

We used microclimate data collected by @Nitta2020.
This dataset includes relative humidity (RH) and temperature measured every 15 min from `r climate_start` to `r climate_end` at the sampling sites with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA).
The dataloggers were outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation, and placed in pairs at each site (one 'epiphytic' datalogger at \circa{ }1.5 m on a tree, one 'terrestrial' datalogger at \circa{ }10 cm next to the tree).
During the study period, some dataloggers malfunctioned, presumably due to prolonged exposure to high humidity.
Any days during which one or more dataloggers failed are excluded from the dataset.
One terrestrial datalogger (Mt. Tohiea 393 m site) was lacking data for more than half of the survey period, but had data available for the same period from a preliminary survey conducted during the previous year (2012--2013), so the previous year's data for this site was used instead.
The dataset includes `r climate_n_days` days of data from `r climate_n_sites` sites (`r climate_n_loggers` dataloggers).

@@

Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8--16 individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment chlorophyll fluorescence (F~v~/F~m~) was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100% RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22°C to 25°C.
Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (-282 MPa), Mg(NO~3~)~2~ (-86 MPa), and NaCl (-38 MPa).
Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible; we therefore used the same treatment for all gametophytes (2 d at -86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described below.
Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
Following the desiccation treatment, plants were rewetted using water from a local stream.
Plants were allowed to rehydrate, and chlorophyll fluorescence was again measured at 0.5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment chlorophyll fluorescence yield [@WatkinsJr.2007c].

@@

In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes (Table 1) including *Callistopteris apiifolia* (C. Presl) Copel., *Crepidomanes minutum* var. 2 and 3, *Hymenophyllum digitatum* (Sw.) Fosberg, *Hymenophyllum javanicum* Spreng., *Hymenophyllum pallidum* (Blume) Ebihara & K.
Iwats., and *Polyphlebium borbonicum* (Bosch) Ebihara & Dubuisson.
All of these have sporophytes that are normally only found in cloud forest (\circa{ }500--1200 m), but gametophytes that are distributed to lower elevations.

Sporophytes of two taxa, *C. minutum* var. 3 and *Hymenophyllum braithwaitei* (Ebihara & K. Iwats.), did not occur in the study plots, but each was observed from a single population on Mt. Tohiea, at \circa{ }1000 m and near the summit (1206 m), respectively.
We include these observations for comparison of ranges between sporophytes and gametophytes.

@@

Gametophytes of low elevation epiphytes had the highest DT (\circa{ }80% recovery), followed by high elevation epiphytes and saxicolous species (both with recovery from \circa{ }25% to 75%), with terrestrial species essentially non-tolerant (less than 20% recovery; Fig. 5).
DT of gametophytes did not significantly exceed that of sporophytes for any species except for saxicolous *P. endlicherianum* (Fig. 5; *t*-test, *P* = `r ttest_pval("Polyphlebium_endlicherianum", "recover")`), and was significantly lower in gametophytes than sporophytes of some high elevation epiphytic species including *C. minutum* var. 2, *H.* *digitatum*, *H. polyanthos*, and *P. borbonicum* (Fig. 5; *t*-test, *P* = `r ttest_pval("Crepidomanes_minutum2", "recover")`, `r ttest_pval("Hymenophyllum_digitatum", "recover")`, `r ttest_pval("Hymenophyllum_polyanthos", "recover")`, and `r ttest_pval("Polyphlebium_borbonicum", "recover")`, respectively).

@@

```{r pgls-stats}
sporo_vpd_pval <- env_range_dt_model_summary %>% filter(model == "sporo_vpd_model") %>% pull(p_value) %>%
  round_t(2)
gameto_vpd_pval <- env_range_dt_model_summary %>% filter(model == "gameto_vpd_model") %>% pull(p_value) %>%
  round_t(2)
```

Desiccation tolerance tended to increase with VPD for both gametophytes and sporophytes (Fig. 6).
These trends were significant when analyzed using PGLS for gametophytes (*P* = `r gameto_vpd_pval`) but not sporophytes (*P* = `r sporo_vpd_pval`; Appendix S5).

@@

We found a wide variation of DT in filmy fern sporophytes on Moorea, from no tolerance in terrestrial species to tolerance of water potentials well below -200 MPa in some epiphytes.
Levels of similarly extreme DT have also been reported in other species of *Hymenophyllum* [@Proctor2003; @Proctor2012] and a general correlation of habitat with DT in filmy ferns has been previously established [@Shreve1911; @Proctor2003; @Proctor2012; @Parra2009; @Saldana2013], but no other studies to our knowledge have documented a similar range of DT values in filmy ferns from a single site.
One may wonder then, what is the adaptive significance of being able to withstand drying at -200 MPa vs. -100 MPa when the ambient water potential is typically > -5 MPa and only rarely, if ever, drops below -50 MPa.
The recent finding of a constitutive mechanism for DT in filmy ferns [@GarcesCea2014] supports the need for a relatively wide "safety margin," and similar (or even more extreme) levels of constitutive DT are well known in bryophytes [@Oliver1993].
The conservative behavior of fern stomata, which lack a response to the plant stress hormone ABA, is well documented [@Brodribb2011; @Mcadam2012; @Mcadam2013].
It is possible that filmy ferns also behave conservatively with respect to water-stress.
We suggest that the variation in DT observed here is relevant because it is correlated with habitat, and likely reflects overall physiological tolerance, not just artificial extremes.
For example, sporophytes of some high elevation epiphytes that we surveyed (*C. minutum* var. 2, *H. digitatum*, *H. multifidum*, *H. polyanthos*) seem to be able to better withstand intense vs. moderate desiccation, with higher recovery values at -282 and -86 MPa vs. -38 MPa when dry for 15 d.
This may be an adaptation to cloud forest conditions; although cloud forests are frequently moist due to constant fog, dry periods are intense during the times when there is no cloud cover.
The complete lack of DT in terrestrial species is not surprising, considering they have much more direct access to water than epiphytes and therefore substantially less selective pressure to maintain DT.
Since controls kept at 100% RH maintained F~v~/F~m~ values near pre-treatment levels in all sporophytes (Appendix S3), it is unlikely that the lack of tolerance seen in terrestrial species is due solely to removal from the soil.
Comparison of DT with the minimum water potential envelope of each species (Fig. 6) indicates that sporophytes of non-tolerant terrestrial and saxicolous species may not be able to occur in habitats with minimum water potential exceeding \circa{ }-7 MPa.
In contrast, habitats with minimum water potential below -7 MPa are nearly all epiphytic, and the filmy fern sporophytes occurring there are all capable of DT.

@@

Contrary to our hypotheses, gametophytes of filmy ferns are in general not more tolerant of the abiotic environment (light levels and water potential) than sporophytes.
This differs from other fern groups with desiccation tolerant gametophytes but desiccation sensitive sporophytes [@WatkinsJr.2007c].
In the filmy ferns we studied, DT levels of both sporophytes and gametophytes largely reflected habitat; in general, epiphytic species tended to have higher DT than terrestrial or saxicolous species, regardless of life stage.
Interestingly, this echoes much earlier field observations (not experimental data) by @Holloway1930 on filmy fern gametophytes in New Zealand: he reported apparent DT in gametophytes of *Hymenophyllum rarum* R. Br. and *H. villosum* Colenso, both high elevation epiphytes with desiccation tolerant sporophytes, and lack of DT in terrestrial filamentous gametophytes of *Polyphlebium colensoi* (Hook.) Ebihara & K. Iwats., *Abrodictyum strictum* (Menzies ex Hook. & Grev.) Ebihara & K. Iwats., and *A. elongatum* (A. Cunn.) Ebihara & K. Iwats. occupying moist, protected sites on the forest floor.
@WatkinsJr.2007c also found that DT of fern gametophytes tended to match habitat, with terrestrial species having lower DT than epiphytes.
However, we found that within high elevation epiphytes, gametophytes had lower DT than sporophytes, and only one species (saxicolous *P. endlicherianum*) with gametophytes that had significantly greater DT than sporophytes (Fig. 5).
At high elevation sites (\ie, cloud forest), epiphytic ferns often occur in a dense matrix of bryophytes (Fig. 1B).
It is possible that this bryophyte cover helps to retain moisture during brief periods of drought and thus protect the epiphytic fern gametophytes that grow amongst them.
More detailed studies of the high elevation epiphytic environment and interactions between co-occurring gametophytes of ferns and bryophytes are needed to test this hypothesis.

Although we predicted that species with gametophytes that occur beyond the range of their sporophytes would have higher DT and be adapted for higher light levels, this was not the case.
This is perhaps best illustrated by *C. apiifolia* (Fig. 1D, E, H).
Sporophytes of this species are usually restricted to moist cloud forest sites from \circa{ }600 m to 1200 m, but we observed gametophytes over a much wider range, down to 200 m.
A similar distribution pattern has been reported in congeneric *Callistopteris baldwinii* (D. C. Eaton) Copel.
from Hawaii [@Dassler1997].
Since gametophytes of *C. apiifolia* are common on Moorea and have a unique morphology [@Nitta2017], we were able to collect enough material of this species for additional tests of DT at higher humidity levels (48 h at 75% and 100% RH; results not shown).
However, it still failed to recover from even these gentle treatments.
This suggests that the widespread gametophytes of *C. apiifolia* do not rely on DT to attain their distribution, but instead must do so via some other mechanism.
Although we measured terrestrial microhabitats by positioning sensors a few cm above the ground, these sensors are still outside of the boundary layer.
It is likely that the gametophytes of *C. apiifolia* do not actually experience the conditions measured with our 'terrestrial' dataloggers, but instead grow within a much more consistently moist boundary layer.
Gametophytes of *C. apiifolia* produce gemmae (Fig. 1E, H), and often form dense clonal mats.
In the case that their microhabitats do become too dry, they may suffer loss at the edge of the population, then recover by clonal growth (Farrar et al., 2008).
Furthermore, the complex surface of clonal mats has been shown to increase the boundary layer and decrease evaporation rates in mosses [@Rice2001], and may function similarly in *C. apiifolia*.
Other filmy fern species also produce gemmae, but they are much more frequent in *C. apiifolia*, and the large size of the clonal mats formed by *C. apiifolia* indicate that it may have a high growth rate (J. Nitta, pers. obs.).
It is possible that there is a tradeoff between growth rate and DT, such that species capable of rapid asexual growth like *C. apiifolia* are unable to tolerate desiccation, but slower growing epiphytes such as *Hymenophyllum* spp.
can better tolerate it [@Oliver2000].
Growth rates of tolerant and non-tolerant gametophytes including quantification of gemmae production should be investigated in these species to test this hypothesis.

@@

Unlike DT, photosynthetic parameters did not vary significantly between species from different habitats.
Rather, we observed a clear intergenerational difference, with gametophytes consistently adapted for lower light levels than sporophytes (Fig. 5, Table 3).
In the only other study to our knowledge that compared photosynthetic rates between gametophytes and sporophytes of a filmy fern, @Johnson2000 also found that gametophytes of *V. speciosa* reached maximum ETR at lower levels than sporophytes: in gametophytes of this species, photosynthesis became saturated at a PPFD of 4--5 µmol·m^-2^·s^-1^ in the field or 30--50 µmol·m^-2^·s^-1^ in the lab vs. 30--50 µmol·m^-2^ ·s^-1^ for sporophytes (lab measurement only).
Although we found similar differences in direction between gametophytes and sporophytes, the measurements of @Johnson2000 are considerably lower than what we observed in the field for both Moorean filmy fern gametophytes (PPFD~95%~ \circa{ }40--60 µmol·m^-2^·s^-1^) and sporophytes (PPFD~95%~ \circa{ }130--160 µmol·m^-2^·s^-1^).
This makes sense in light of the fact that *V. speciosa* is an extremely deep-shade plant, with gametophytes occupying protected rock crevices that receive less than 1 µmol·m^-2^·s^-1^ PPFD for most of the day [@Johnson2000]; many of the gametophytes sampled in our study occur in more exposed conditions.
It is surprising, however, that we did not observe any differences between species in different habitats, given that light levels should vary between tree trunks and the forest floor, at least on a \circa{ }1 m scale.
It is possible that the actual light levels experienced by gametophytes growing in these habitats may not vary significantly if they occupy protected microsites such as bryophyte mats or crevices in rock or bark.

@@

Previous studies of filmy fern ecophysiology also found a link between habitat and DT and/or photosynthetic rates [@Proctor2003; @Proctor2012; @Parra2009; @Saldana2013]; however, ours is the first to our knowledge to investigate this pattern while taking phylogeny into account.
We detected significant phylogenetic signal in DT, suggesting that phylogeny should be considered when interpreting differences in DT between species (Table 2).
In contrast, the lack of phylogenetic signal in photosynthetic traits, coupled with the clear intergenerational differences we observed, suggest that sporophytes and gametophytes are adapted for different light levels regardless of clade or ecological niche.
In a more densely sampled study of morphological evolution in the trichomanoid clade, @Dubuisson2013 identified several morphological changes including reduced stele and root systems associated with the evolution of epiphytic growth within the HE subclade (comprising the genera *Polyphlebium*, *Didymoglossum*, *Crepidomanes*, and *Vandenboschia*).
It is possible that both reduced morphology and increased DT are parts of an integrated adaptive strategy for epiphytic growth within this subclade of filmy ferns, and have evolved together repeatedly.
However, @Dubuisson2013 were unable to determine if these morphological changes had occurred independently, or if there was a single origin at the base of the HE clade followed by multiple losses.
Increased sampling of both morphological and physiological traits of additional trichomanoid taxa is needed to distinguish between these scenarios.

@@

In filmy ferns, both sporophyte and gametophyte generations are poikilohydric; however, we have found here that this does not mean they necessarily share similar physiological optima.
Although filmy fern sporophytes are often of small stature, elevating a leaf even a few cm above the substrate surface means escape from the boundary layer and exposure to substantially greater evapotranspiration.
It is likely that even in this clade, wherein the physiological divide between sporophyte and gametophyte is probably the smallest amongst ferns, the two generations experience substantially different selective pressures.
Fossils indicate that some of the earliest vascular plants may have had sporophytes and gametophytes that were both branched and small [@Kenrick1997a], but it is unclear at what point selective pressures began to diverge to favor large, complex sporophytes and small, simple gametophytes.
The relevance of plant size to water relations is supported by a study showing increased DT in smaller size classes of the epiphytic fern *Asplenium auritum* Sw. [@Testo2012].
Future studies of DT in filmy ferns should include juvenile stages of the sporophyte, which are expected to experience similar microhabitats as gametophytes, to better determine the role of microhabitat vs. life stage (and plant size) on DT.
Finally, more detailed analyses of the strategies used by each generation to cope with desiccation stress, such as transcriptomics, are needed to better understand the ecology and evolution of filmy ferns as well as the transition from desiccation tolerance to avoidance in vascular plants.

@@

@@

<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
